---
title: "Pest Control"
author: "A Case Study in Bayesian Workflow"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
---

## Setup

```{r setup}
knitr::opts_chunk$set(
  echo = TRUE, 
  dev = "png",
  dpi = 150,
  fig.align = "center",
  comment = NA
)
library(rstan)
library(dplyr)
library(lubridate)
library(ggplot2)
library(bayesplot)

theme_set(bayesplot::theme_default())

# seed for R's pseudo-RNGs, not Stan's
set.seed(1123) 
```

## The problem

### Background

Imagine that you are a statistician or data scientist working as an independent
contractor. One of your clients is a company that owns many residential buildings 
throughout New York City. The property manager explains that they are concerned about the number
of cockroach complaints that they receive from their buildings. Previously
the company has offered monthly visits from a pest inspector as a solution to
this problem. While this is the default solution of many property managers in
NYC, the tenants are rarely home when the inspector visits, and so the manager
reasons that this is a relatively expensive solution that is currently not very
effective.

One alternative to this problem is to deploy long term bait stations. In this
alternative, child and pet safe bait stations are installed throughout the
apartment building. Cockroaches obtain quick acting poison from these stations
and distribute it throughout the colony. The manufacturer of these bait stations
provides some indication of the space-to-bait efficacy, but the manager suspects
that this guidance was not calculated with NYC roaches in mind. NYC roaches, the
manager rationalizes, have more hustle than traditional roaches; and NYC
buildings are built differently than other common residential buildings in the
US. This is particularly important as the unit cost for each bait station per
year is quite high.

### The goal

The manager wishes to employ your services to help them to find the optimal
number of roach bait stations they should place in each of their buildings in
order to minimize the number of cockroach complaints while also keeping
expenditure on pest control affordable. 

A subset of the company's buildings have been randomly selected for an experiment: 

* At the beginning of each month, a pest inspector randomly places a number of
bait stations throughout the building, without knowledge of the current
cockroach levels in the building
* At the end of the month, the manager records
the total number of cockroach complaints in that building. 
* The manager would like to determine the optimal number of traps ($\textrm{traps}$) that
balances the lost revenue ($R$) that complaints ($\textrm{complaints}$) generate
with the all-in cost of maintaining the traps ($\textrm{TC}$). 

Fortunately, Bayesian data analysis provides a coherent framework for us to tackle this problem.

Formally, we are interested in finding

$$
\arg\max_{\textrm{traps} \in \mathbb{N}} \mathbb{E}_{\text{complaints}}[R(\textrm{complaints}(\textrm{traps})) - \textrm{TC}(\textrm{traps})]
$$

The property manager would also, if possible, like to learn how these results 
generalize to buildings they haven't treated so they can understand the
potential costs of pest control at buildings they are aquiring as well as for
the rest of their building portfolio.

As the property manager has complete control over the number of traps set, the
random variable contributing to this expectation is the number of complaints
given the number of traps. We will model the number of complaints as a function
of the number of traps.

## The data

The data provided to us is in a file called `pest_data.RDS`. Let's
load the data and see what the structure is:

```{r load-data}
pest_data <- readRDS('data/pest_data.RDS')
str(pest_data)
```

We have access to the following fields: 

* `complaints`: Number of complaints per building per month
* `building_id`: The unique building identifier
* `traps`: The number of traps used per month per building
* `date`: The date at which the number of complaints are recorded
* `live_in_super`: An indicator for whether the building as a live-in super
* `age_of_building`: The age of the building
* `total_sq_foot`: The total square footage of the building
* `average_tenant_age`: The average age of the tenants per building
* `monthly_average_rent`: The average monthly rent per building
* `floors`: The number of floors per building

First, let's see how many buildings we have data for:

```{r describe-data}
N_buildings <- length(unique(pest_data$building_id))
N_buildings
```

And make some plots of the raw data: 

```{r data-plots}
ggplot(pest_data, aes(x = complaints)) + 
  geom_bar()

ggplot(pest_data, aes(x = traps, y = complaints, color = live_in_super == TRUE)) + 
  geom_jitter()
```

```{r, data-plots-ts, fig.width = 6, fig.height = 8}
ggplot(pest_data, aes(x = date, y = complaints, color = live_in_super == TRUE)) + 
  geom_line(aes(linetype = "Number of complaints")) + 
  geom_point(color = "black") + 
  geom_line(aes(y = traps, linetype = "Number of traps"), color = "black", size = 0.25) + 
  facet_wrap(~building_id, scales = "free", ncol = 2, labeller = label_both) + 
  scale_x_date(name = "Month", date_labels = "%b") + 
  scale_y_continuous(name = "", limits = range(pest_data$complaints)) + 
  scale_linetype_discrete(name = "") + 
  scale_color_discrete(name = "Live-in super")
```


The first question we'll look at is just whether the number of complaints per
building per month is associated with the number of bait stations per building
per month, ignoring the temporal and across-building variation (we'll come back
to those sources of variation later in the document). That requires only two
variables, $\textrm{complaints}$ and $\textrm{traps}$. How should we model the
number of complaints?


## Bayesian workflow

See slides

## Modeling count data : Poisson distribution

We already know some rudimentary information about what we should expect. The
number of complaints over a month should be either zero or an integer. The
property manager tells us that it is possible but unlikely that number of
complaints in a given month is zero. Occasionally there are a very large number
of complaints in a single month. A common way of modeling this sort of skewed,
single bounded count data is as a Poisson random variable. One concern about
modeling the outcome variable as Poisson is that the data may be
over-dispersed, but we'll start with the Poisson model and then check 
whether over-dispersion is a problem by comparing our model's predictions 
to the data.

### Model 

Given that we have chosen a Poisson regression, we define the likelihood to be
the Poisson probability mass function over the number bait stations placed in
the building, denoted below as `traps`. This model assumes that the mean and
variance of the outcome variable `complaints` (number of complaints) is the
same. We'll investigate whether this is a good assumption after we fit the 
model.

For building $b = 1,\dots,10$ at time (month) $t = 1,\dots,12$, we have

$$
\begin{align*}
\textrm{complaints}_{b,t} & \sim \textrm{Poisson}(\lambda_{b,t}) \\
\lambda_{b,t} & = \exp{(\eta_{b,t})} \\
\eta_{b,t} &= \alpha + \beta \, \textrm{traps}_{b,t}
\end{align*}
$$

Let's encode this probability model in a Stan program. 

### Writing our first Stan model

* Write `simple_poisson_regression.stan` together

### Making sure our code is right

However, before we fit the model to real data, we should check that our
model works well with simulated data. We'll simulate data according to the model
and then check that we can sufficiently recover the parameter values used
in the simulation.

* Write `simple_poisson_regression_dgp.stan` together

We can use the `stan()` function to compile and fit the model, but here we will 
do the compilation and fitting in two stages to demonstrate what is really 
happening under the hood. 

First we will compile the Stan program (`simple_poisson_regression_dgp.stan`) that
will generate the fake data.

```{r , cache=TRUE, results="hide", message=FALSE}
comp_dgp_simple <- stan_model('stan_programs/simple_poisson_regression_dgp.stan')
```

Printing this object shows the Stan program: 

```{r}
print(comp_dgp_simple)
```

Now we can simulate the data by calling the `sampling()` function.

```{r runpoissondgp}
fitted_model_dgp <- sampling(
  comp_dgp_simple,
  data = list(N = nrow(pest_data), mean_traps = mean(pest_data$traps)),
  chains = 1,
  iter = 1,
  algorithm = 'Fixed_param',
  seed = 123
  )

# see http://mc-stan.org/rstan/articles/stanfit_objects.html for various
# ways of extracting the contents of the stanfit object
samps_dgp <- rstan::extract(fitted_model_dgp)
str(samps_dgp)
```


### Fit the model to the fake data:

In order to pass the fake data to our Stan program using RStan, we need to
arrange the data into a named list. The names must match the names used 
in the `data` block of the Stan program.

```{r}
stan_dat_fake <- list(
  N = nrow(pest_data), 
  traps = samps_dgp$traps[1, ], 
  complaints = samps_dgp$complaints[1, ]
)
str(stan_dat_fake)
```

Now that we have the simulated data we fit the model to see if we can recover
the `alpha` and `beta` parameters used in the simulation.

```{r , cache=TRUE, results="hide", message=FALSE}
comp_model_P <- stan_model('stan_programs/simple_poisson_regression.stan')
fit_model_P <- sampling(comp_model_P, data = stan_dat_fake, seed = 123)

posterior_alpha_beta <- as.matrix(fit_model_P, pars = c('alpha','beta'))
head(posterior_alpha_beta)
```


### Assess parameter recovery

```{r}
true_alpha_beta <- c(samps_dgp$alpha, samps_dgp$beta)
mcmc_recover_hist(posterior_alpha_beta, true = true_alpha_beta)
```

It's possible we dont' do a great job recovering the parameters here simply
because we're simulating so few observations that the posterior uncertainty
remains rather large, but it looks at least plausible. 
If we did the simulation with many more observations the
parameters would be estimated much more precisely.

We should also check if the `y_rep` datasets (in-sample predictions) that we coded 
in the `generated quantities` block are similar to the `y` (complaints) values
we conditioned on when fitting the model. (The **bayesplot**
package [vignettes](http://mc-stan.org/bayesplot/articles/graphical-ppcs.html)
are a good resource on this topic.)

Here is a plot of the density estimate of the observed data compared to 
200 of the `y_rep` datasets: 
```{r}
ppc_dens_overlay(y = stan_dat_fake$complaints, yrep = y_rep[1:200, ])
```


One common check for regression models is to plot the residuals against the
observed values for $y$. While this plot is what we would expect for a Poisson
regression, it is relatively hard to interpret:

```{r}
y_rep <- as.matrix(fit_model_P, pars = "y_rep")
dim(y_rep)

mean_y_rep <- colMeans(y_rep)
std_resid <- (stan_dat_fake$complaints - mean_y_rep) / sqrt(mean_y_rep)
qplot(mean_y_rep, std_resid) + hline_at(2) + hline_at(-2)
```

One alternative to this is a rootogram. This is a plot of the expected counts
(continuous line) vs the observed counts (blue histogram). We can see the model
fits well because the observed histogram matched the expected counts relatively
well.

```{r}
ppc_rootogram(stan_dat_fake$complaints, yrep = y_rep)
```


### Fit with real data 

To fit the model to real data we'll first create a list to pass to Stan using
the variables in the `pest_data` data frame:

```{r stan-data}
stan_dat_simple <- list(
  N = nrow(pest_data), 
  complaints = pest_data$complaints,
  traps = pest_data$traps
)
```

As we have already compiled the model, we can jump straight to sampling from it.

```{r fit_P_real_data, cache=TRUE}
fit_P_real_data <- sampling(comp_model_P, data = stan_dat_simple)
```

and printing the parameters. What do these tell us? 

```{r results_simple_P}
print(fit_P_real_data, pars = c('alpha','beta'))
```

We can also plot the posterior distributions: 

```{r hist_simple_P}
mcmc_hist(as.matrix(fit_P_real_data, pars = c('alpha','beta')))
```

As we expected, it appears the number of bait stations set in a building is
associated with the number of complaints about cockroaches that were made in the
following month. However, we still need to consider how well the model fits.


### Posterior predictive checking

```{r}
y_rep <- as.matrix(fit_P_real_data, pars = "y_rep")
```

```{r marginal_PPC}
ppc_dens_overlay(y = stan_dat_simple$complaints, y_rep[1:200,])
```

This is a plot of the kernel density estimates of the observed data ($y$,
thicker curve) and 200 simulated data sets ($y_{rep}$, thin curves) from the
posterior predictive distribution. If the model was fitting the data well, there
would be little difference between the observed density and the simulated
density. However, as you can see here, the simulated density is not as dispersed
as the kernel density of the observed data, and doesn't seem to capture the rate
of zeros in the observed data. This indicates that the Poisson model may not be
the best model to use for this data.

Let's explore this further by looking directly at the proportion of zeros in the
real data and predicted data.
```{r}
prop_zero <- function(x) mean(x == 0)
ppc_stat(y = stan_dat_simple$complaints, yrep = y_rep, stat = "prop_zero")
```
The plot above shows the observed proportion of zeros (thick vertical line) and
a histogram of the proportion of zeros in each of the simulated data sets. It is
clear that the model does not capture this feature of the data well at all.

This next plot is a plot of the standardised residuals of the observed vs predicted number of complaints. 

```{r}
mean_y_rep <- colMeans(y_rep)
std_resid <- (stan_dat_simple$complaints - mean_y_rep) / sqrt(mean_y_rep)
qplot(mean_y_rep, std_resid) + hline_at(2) + hline_at(-2)
```

As you can see here, it looks as though we have more positive residuals than negative,
which indicates that the model tends to underestimate the number of complaints
that will be received.

The rootogram is another useful plot to compare the observed vs expected number
of complaints. This is a plot of the expected counts (continuous line) vs the
observed counts (blue histogram):

```{r}
ppc_rootogram(stan_dat_simple$complaints, yrep = y_rep)
```

If the model was fitting well these would be relatively similar, however in this
figure we can see the number of complaints is underestimated if there are few
complaints, over-estimated for medium numbers of complaints and underestimated
if there are a large number of complaints.

We can also view how the predicted number of complaints varies with the number
of traps. From this we can see that the model doesn't seem to fully capture the
data.

```{r}
ppc_intervals(
  y = stan_dat_simple$complaints, 
  yrep = y_rep,
  x = stan_dat_simple$traps
) + 
  labs(x = "Number of traps", y = "Number of complaints")
```

Specifically, the model doesn't capture the tails of the observed data very
well.


## Expanding the model: multiple predictors

Modeling the relationship between complaints and bait stations is the simplest 
model. We can expand the model, however, in a few ways that will be beneficial 
for our client. Moreover, the manager has told us that they expect there are a
number of other reasons that one building might have more roach complaints than
another.

### Interpretability

Currently, our model's mean parameter is a rate of complaints per 30 days, but
we're modeling a process that occurs over an area as well as over time. We have
the square footage of each building, so if we add that information into the
model, we can interpret our parameters as a rate of complaints per square foot
per 30 days.

$$
\begin{align*}
\textrm{complaints}_{b,t} & \sim \textrm{Poisson}(\textrm{sq_foot}_b\,\lambda_{b,t}) \\
\lambda_{b,t} & = \exp{(\eta_{b,t} )} \\
\eta_{b,t} &= \alpha + \beta \, \textrm{traps}_{b,t}
\end{align*}
$$

The term $\text{sq_foot}$ is called an exposure term. If we log the term, we can 
put it in $\eta_{b,t}$:

$$
\begin{align*}
\textrm{complaints}_{b,t} & \sim \textrm{Poisson}(\lambda_{b,t}) \\
\lambda_{b,t} & = \exp{(\eta_{b,t} )} \\
\eta_{b,t} &= \alpha + \beta \, \textrm{traps}_{b,t} + \textrm{log_sq_foot}_b
\end{align*}
$$

A quick test shows us that there appears to be a relationship between the square
footage of the building and the number of complaints received:

```{r}
ggplot(pest_data, aes(x = log(total_sq_foot), y = log1p(complaints))) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```

Using the property manager's intuition, we include two extra peices of
information we know about the building - the (log of the) square floor space and
whether there is a live in super or not - into both the simulated and real data.
 
```{r}
stan_dat_simple$log_sq_foot <- log(pest_data$total_sq_foot/1e4)
stan_dat_simple$live_in_super <- pest_data$live_in_super
```


### Stan program for Poisson multiple regression

* Write `multiple_poisson_regression.stan` together

### Simlute fake data with multiple predictors

Now we need a new Stan model that uses multiple predictorst. Firs we have to
compile `multiple_poisson_regression_dgp.stan`.

```{r compmultPDGP, cache=TRUE, results="hide", message=FALSE}
comp_dgp_multiple <- stan_model('stan_programs/multiple_poisson_regression_dgp.stan')
```

Then we generate some fake data from this model. 

```{r runpoissondgp2}
fitted_model_dgp <-
  sampling(
  comp_dgp_multiple,
  data = list(N = nrow(pest_data)),
  chains = 1,
  cores = 1,
  iter = 1,
  algorithm = 'Fixed_param',
  seed = 123
  )
samps_dgp <- rstan::extract(fitted_model_dgp)
```

Now pop that data into a list ready for Stan. 

```{r}
stan_dat_fake <- list(
  N = nrow(pest_data), 
  log_sq_foot = samps_dgp$log_sq_foot[1, ],
  live_in_super = samps_dgp$live_in_super[1, ],
  traps = samps_dgp$traps[1, ], 
  complaints = samps_dgp$complaints[1, ]
)
```

And then compile and fit the model we wrote for the multiple regression.

```{r, cache=TRUE, message=FALSE, warning=FALSE}
comp_model_P_mult <- stan_model('stan_programs/multiple_poisson_regression.stan')
fit_model_P_mult <- sampling(comp_model_P_mult, data = stan_dat_fake, chains = 4, cores = 4)
```

Then compare these parameters to the true parameters:

```{r}
posterior_alpha_beta <- as.matrix(fit_model_P_mult, pars = c('alpha','beta','beta_super'))
true_alpha_beta <- c(samps_dgp$alpha,samps_dgp$beta,samps_dgp$beta_super)
mcmc_recover_hist(posterior_alpha_beta, true = true_alpha_beta)
```

We've recovered the parameters well, so we've probably coded the Stan program
correctly and we're ready to fit the real data.

### Fit the real data

Now let's use the real data and explore the fit.

```{r fit_mult_P_real_dat}
fit_model_P_mult_real <- sampling(comp_model_P_mult, data = stan_dat_simple)
y_rep <- as.matrix(fit_model_P_mult_real, pars = "y_rep")
ppc_dens_overlay(stan_dat_simple$complaints, y_rep[1:200,])
```

This again looks like we haven't captured the smaller counts very well, nor
have we captured the larger counts.


```{r}
prop_zero <- function(x) mean(x == 0)
ppc_stat(y = stan_dat_simple$complaints, yrep = y_rep, stat = "prop_zero", binwidth = 0.01)
```

We're still severely underestimating the proportion of zeros in the data. Ideally
this vertical line would fall somewhere within the histogram.

We can also plot uncertainty intervals for the predicted complaints for different
numbers of traps.

```{r}
ppc_intervals(
  y = stan_dat_simple$complaints, 
  yrep = y_rep,
  x = stan_dat_simple$traps
) + 
  labs(x = "Number of traps", y = "Number of complaints")
```

We can see that we've increased the tails a bit more at the larger numbers of traps
but we still have some large observed numbers of complaints that the model
would consider extremely unlikely events. 
 
## Modeling count data with the Negative Binomial

When we considered modelling the data using a Poisson, we saw that the model
didn't appear to fit as well to the data as we would like. In particular the
model underpredicted low and high numbers of complaints, and overpredicted the
medium number of complaints. This is one indication of over-dispersion, where
the variance is larger than the mean. A Poisson model doesn't fit over-dispersed
count data very well because the same parameter $\lambda$, controls both the
expected counts and the variance of these counts. The natural alternative to
this is the negative binomial model:

$$
\begin{align*}
\text{complaints}_{b,t} & \sim \text{Neg-Binomial}(\lambda_{b,t}, \phi) \\
\lambda_{b,t} & = \exp{(\eta_{b,t})} \\
\eta_{b,t} &= \alpha + \beta \, {\rm traps}_{b,t} + \beta_{\rm super} \, {\rm super}_{b} + \text{log_sq_foot}_{b}
\end{align*}
$$

In Stan the negative binomial mass function we'll use is called 
$\texttt{neg_binomial_2_log}(\text{ints} \, y, \text{reals} \, \eta, \text{reals} \, \phi)$ 
in Stan. Like the `poisson_log` function, this negative binomial mass function
that is parameterized in terms of its log-mean, $\eta$, but it also has a
precision $\phi$ such that

$$
\mathbb{E}[y] \, = \lambda = \exp(\eta)
$$

$$
\text{Var}[y] = \lambda + \lambda^2/\phi = \exp(\eta) + \exp(\eta)^2 / \phi.
$$ 

As $\phi$ gets larger the term $\lambda^2 / \phi$ approaches zero and so 
the variance of the negative-binomial approaches $\lambda$, i.e., the
negative-binomial gets closer and closer to the Poisson.

### Stan program for negative-binomial regression

* Write `multiple_NB_regression.stan` together


### Fake data fit: Multiple NB regression

```{r , cache=TRUE, results="hide", message=FALSE}
comp_dgp_multiple_NB <- stan_model('stan_programs/multiple_NB_regression_dgp.stan')
```

We're going to generate one draw from the fake data model so we can use the data
to fit our model and compare the known values of the parameters to the posterior
density of the parameters.

```{r fake_data_dgp_NB}
fitted_model_dgp_NB <-
  sampling(
  comp_dgp_multiple_NB,
  data = list(N = nrow(pest_data)),
  chains = 1,
  cores = 1,
  iter = 1,
  algorithm = 'Fixed_param',
  seed = 123
  )
samps_dgp_NB <- rstan::extract(fitted_model_dgp_NB)
```

Create a dataset to feed into the Stan model.

```{r NB_fake_stan_dat}
stan_dat_fake_NB <- list(
  N = nrow(pest_data), 
  log_sq_foot = samps_dgp_NB$log_sq_foot[1, ],
  live_in_super = samps_dgp_NB$live_in_super[1, ],
  traps = samps_dgp_NB$traps[1, ], 
  complaints = samps_dgp_NB$complaints[1, ]
)
```

Compile the inferential model.

```{r , cache=TRUE, results="hide", message=FALSE}
comp_model_NB <- stan_model('stan_programs/multiple_NB_regression.stan')
```

Now we run our NB regression over the fake data and extract the samples to
examine posterior predictive checks and to check whether we've sufficiently
recovered our known parameters, $\text{alpha}$ $\texttt{beta}$, .

```{r runNBoverfake, message=FALSE, warning=FALSE}
fitted_model_NB <- sampling(comp_model_NB, data = stan_dat_fake_NB, 
                            chains = 4, cores = 4)
posterior_alpha_beta_NB <- 
  as.matrix(fitted_model_NB,
            pars = c('alpha',
                     'beta',
                     'beta_super',
                     'inv_phi')
  )
```

Construct the vector of true values from your simulated dataset and compare to
the recovered parameters.
```{r}
true_alpha_beta_NB <- 
  c(samps_dgp_NB$alpha,
    samps_dgp_NB$beta,
    samps_dgp_NB$beta_super,
    samps_dgp_NB$inv_phi
  )
mcmc_recover_hist(posterior_alpha_beta_NB, true = true_alpha_beta_NB)
```


### Fit to real data and check the fit

```{r runNB}
fitted_model_NB <- sampling(comp_model_NB, data = stan_dat_simple)
samps_NB <- rstan::extract(fitted_model_NB)
```

Let's look at our predictions vs. the data.

```{r ppc-full}
y_rep <- samps_NB$y_rep
ppc_dens_overlay(stan_dat_simple$complaints, y_rep[1:200,])
```

It appears that our model now captures both the number of small counts better
as well as the tails. 

Let's check if the negative binomial model does a better job capturing
the number of zeros: 

```{r}
ppc_stat(y = stan_dat_simple$complaints, yrep = y_rep, stat = "prop_zero")
```

These look OK, but let's look at the standardized residual plot.

```{r}
mean_inv_phi <- mean(samps_NB$inv_phi)
mean_y_rep <- colMeans(y_rep)
std_resid <- (stan_dat_simple$complaints - mean_y_rep) / sqrt(mean_y_rep + mean_y_rep^2*mean_inv_phi)
qplot(mean_y_rep, std_resid) + hline_at(2) + hline_at(-2)
```

Looks ok, but we still have some large standardized residuals. This might be
because we are currently ignoring that the data are clustered by buildings, and
that the probability of roach issue may vary substantially across buildings.

```{r}
ppc_rootogram(stan_dat_simple$complaints, yrep = y_rep)
```

The rootgram now looks much more plausible. We can tell this because now the
expected number of complaints matches much closer to the observed number of
complaints. However, we still have some larger counts that appear
to be outliers for the model.

Check predictions by number of traps: 
```{r}
ppc_intervals(
  y = stan_dat_simple$complaints, 
  yrep = y_rep,
  x = stan_dat_simple$traps
) + 
  labs(x = "Number of traps", y = "Number of complaints")
```

We haven't used the fact that the data are clustered by building yet. A posterior 
predictive check might elucidate whether it would be a good idea to add the building
information into the model.

```{r ppc-group_means}
ppc_stat_grouped(
  y = stan_dat_simple$complaints, 
  yrep = y_rep, 
  group = pest_data$building_id, 
  stat = 'mean', 
  binwidth = 0.5
)
```

We're getting plausible predictions for most building means but some are
estimated better than others and some have very large uncertainties. If we
explicitly model the variation across buildings we may be able to get 
much better estimates. 

## Hierarchical modeling

### Modeling varying intercepts for each building

Let's add a hierarchical intercept parameter, $\alpha_b$ at the building level
to our model.

$$
\text{complaints}_{b,t} \sim \text{Neg-Binomial}(\lambda_{b,t}, \phi) \\
\lambda_{b,t}  = \exp{(\eta_{b,t})} \\
\eta_{b,t} = \mu_b + \beta \, {\rm traps}_{b,t} + \beta_{\rm super}\, {\rm super}_b + \text{log_sq_foot}_b \\
\mu_b \sim \text{Normal}(\alpha, \sigma_{\mu})
$$

In our Stan model, $\mu_b$ is the $b$-th element of the vector
$\texttt{mu}$ which has one element per building.

One of our predictors varies only by building, so we can rewrite the above model 
more efficiently like so:

$$
\eta_{b,t} = \mu_b + \beta \, {\rm traps}_{b,t} + \text{log_sq_foot}_b\\
\mu_b \sim \text{Normal}(\alpha +  \beta_{\text{super}} \, \text{super}_b , \sigma_{\mu})
$$


We have more information at the building level as well, like the average age of
the residents, the average age of the buildings, and the average per-apartment
monthly rent so we can add that data into a matrix called
`building_data`, which will have one row per building and four columns:

* `live_in_super`
* `age_of_building`
* `average_tentant_age`
* `monthly_average_rent`

We'll write the Stan model like:

$$
\eta_{b,t} = \alpha_b + \beta \, {\rm traps} + \text{log_sq_foot}\\
\mu \sim \text{Normal}(\alpha + \texttt{building_data} \, \zeta, \,\sigma_{\mu})
$$

### Prepare building data for hierarchical 

We'll need to do some more data prep before we can fit our models. Firstly to
use the building variable in Stan we will need to trasnform it from a factor
variable to an interger variable.

```{r prep-data}
N_months <- length(unique(pest_data$date))

# Add some IDs for building and month
pest_data <- pest_data %>%
  mutate(
    building_fac = factor(building_id, levels = unique(building_id)),
    building_idx = as.integer(building_fac),
    ids = rep(1:N_months, N_buildings),
    mo_idx = lubridate::month(date)
  )

# Center and rescale the building specific data
building_data <- pest_data %>%
    select(
      building_idx, 
      live_in_super,
      age_of_building,
      total_sq_foot, 
      average_tenant_age, 
      monthly_average_rent
    ) %>% 
    unique() %>%
    arrange(building_idx) %>% 
    select(-building_idx) %>% 
    scale(scale=FALSE) %>%
    as.data.frame() %>%
    mutate( # scale by constants
      age_of_building = age_of_building / 10,
      total_sq_foot = total_sq_foot / 10000,
      average_tenant_age = average_tenant_age / 10,
      monthly_average_rent = monthly_average_rent / 1000
    ) %>%
    as.matrix()

# Make data list for Stan
stan_dat_hier <- 
  with(pest_data, 
        list(complaints = complaints,
             traps = traps,
             N = length(traps),
             J = N_buildings,
             M = N_months,
             log_sq_foot = log(pest_data$total_sq_foot/1e4),
             building_data = building_data[,-3],
             mo_idx = as.integer(as.factor(date)),
             K = 4,
             building_idx = building_idx
             )
        )
```

### Compile and fit the hierarchical model

Let's compile the model. 

```{r comp-NB-hier, cache=TRUE, results="hide", message=FALSE}
comp_model_NB_hier <- stan_model('stan_programs/hier_NB_regression.stan')
```

Fit the model to data.

```{r run-NB-hier}
fitted_model_NB_hier <-
  sampling(
  comp_model_NB_hier,
  data = stan_dat_hier,
  chains = 4,
  cores = 4,
  iter = 4000
  )
```

### Diagnostics 

We get a bunch of warnings from Stan about divergent transitions, which is 
an indication that there may be regions of the posterior that have not been
explored by the Markov chains.

Divergences are discussed in more detail in the course slides as well as 
the **bayesplot** (MCMC diagnostics vignette)[http://mc-stan.org/bayesplot/articles/visual-mcmc-diagnostics.html]
and [_A Conceptual Introduction to Hamiltonian Monte Carlo_](https://arxiv.org/abs/1701.02434).

In this
example we will see that we have divergent transitions because we need to
reparameterize our model - i.e., we will retain the overall structure of the
model, but transform some of the parameters so that it is easier for Stan to
sample from the parameter space. Before we go through exactly how to do this
reparameterization, we will first go through what indicates that this is
something that reparamerization will resolve. We will go through:

1. Examining the fitted parameter values, including the effective sample size
2. Traceplots and scatterplots that reveal particular patterns in locations of the divergences.

First let's extract the fits from the model. 

```{r} 
samps_hier_NB <- rstan::extract(fitted_model_NB_hier)
```

Then we print the fits for the parameters that are of most interest. 

```{r print-NB-hier}
print(fitted_model_NB_hier, pars = c('sigma_mu','beta','alpha','phi','mu'))
```
You can see that the effective samples are quite low for many of the parameters
relative to the total number of samples. This alone isn't indicative of the need
to reparameterize, but it indicates that we should look further at the trace
plots and pairs plots. First let's look at the traceplots to see if the divergent
transistions form a pattern.

```{r}
# use as.array to keep the markov chains separate for trace plots
mcmc_trace(
  as.array(fitted_model_NB_hier,pars = 'sigma_mu'), 
  np = nuts_params(fitted_model_NB_hier),
  window = c(500,1000)
)
```

Looks as if the divergent parameters, the little red bars underneath the traceplots
correspond to samples where the sampler gets stuck at one parameter value for $\sigma_\mu$.

```{r}
# assign to object so we can compare to another plot later
scatter_with_divs <- mcmc_scatter(
  as.array(fitted_model_NB_hier),
  pars = c("mu[4]", 'sigma_mu'), 
  transform = list('sigma_mu' = "log"), 
  np = nuts_params(fitted_model_NB_hier)
)
scatter_with_divs
```

What we have here is a cloud-like shape, with most of the divergences clustering
towards the bottom. We'll see a bit later that we actually want this to look more
like a funnel than a cloud, but the divergences are indicating that the sampler
can't explore the narrowing neck of the funnel. 

One way to see why we should expect some version of a funnel is to look at some 
simulations from the prior, which we can do without MCMC and thus with no risk 
of sampling problems: 

```{r}
N_sims <- 1000
log_sigma <- rep(NA, N_sims)
theta <- rep(NA, N_sims)
for (j in 1:N_sims) {
  log_sigma[j] <- rnorm(1, mean = 0, sd = 1)
  theta[j] <- rnorm(1, mean = 0, sd = exp(log_sigma[j]))
}
draws <- cbind("mu" = theta, "log(sigma_mu)" = log_sigma)
mcmc_scatter(draws)
```

Of course, if the data is at all informative we shouldn't expect the posterior
to look exactly like the prior. But unless the data is incredibly informative
about the parameters and the posterior concentrates away from the narrow neck of
the funnel, the sampler is going to have to confront the funnel geometry. (See
the [Visual MCMC
Diagnostics](http://mc-stan.org/bayesplot/articles/visual-mcmc-diagnostics.html)
vignette for more on this.)

Another way to look at the divergences is via a parallel coordinates plot: 

```{r}
parcoord_with_divs <- mcmc_parcoord(
  as.array(fitted_model_NB_hier, pars = c("sigma_mu", "mu")),
  np = nuts_params(fitted_model_NB_hier)
)
parcoord_with_divs
```

Again, we see evidence that our problems concentrate when $\texttt{sigma_mu}$ is small.

### Reparameterize and recheck diagnostics

Instead, we should use the non-centered parameterization for $\mu_b$. We
define a vector of auxiliary variables in the parameters block,
$\texttt{mu_raw}$ that is given a $\text{Normal}(0, 1)$ prior in the model
block. We then make $\texttt{mu}$ a transformed parameter:
We can reparameterize the random intercept $\mu_b$, which is distributed:

$$
\mu_b \sim \text{Normal}(\alpha + \texttt{building_data} \, \zeta, \sigma_{\mu})
$$



```
transformed parameters {
  vector[J] mu;
  mu = alpha + building_data * zeta + sigma_mu * mu_raw;
}
```

This gives $\texttt{mu}$ a 
$\text{Normal}(\alpha + \texttt{building_data}\, \zeta, \sigma_\mu)$ 
distribution, but it decouples the dependence of the density of each element of
$\texttt{mu}$ from $\texttt{sigma_mu}$ ($\sigma_\mu$).
hier_NB_regression_ncp.stan uses the non-centered parameterization for
$\texttt{mu}$. We will examine the effective sample size of the fitted model to
see whether we've fixed the problem with our reparameterization.

Compile the model.

```{r comp-NB-hier-ncp, cache=TRUE}
comp_model_NB_hier_ncp <- stan_model('stan_programs/hier_NB_regression_ncp.stan')
```

Fit the model to the data.

```{r run-NB-hier-ncp}
fitted_model_NB_hier_ncp <- sampling(comp_model_NB_hier_ncp, data = stan_dat_hier, chains = 4, cores = 4, control = list(adapt_delta = 0.95))
```

Examining the fit of the new model

```{r n-eff-NB-hier-ncp-check}
print(fitted_model_NB_hier_ncp, pars = c('sigma_mu','beta','alpha','phi','mu'))
```

This has improved the effective sample sizes of $\texttt{mu}$. We extract
the parameters to run our usual posterior predictive checks.

```{r}
scatter_no_divs <- mcmc_scatter(
  as.array(fitted_model_NB_hier_ncp),
  pars = c("mu[4]", 'sigma_mu'), 
  transform = list('sigma_mu' = "log"), 
  np = nuts_params(fitted_model_NB_hier_ncp)
)
bayesplot_grid(scatter_with_divs, scatter_no_divs, 
               grid_args = list(ncol = 2), ylim = c(-11, 1))
```

```{r}
parcoord_no_divs <- mcmc_parcoord(
  as.array(fitted_model_NB_hier_ncp, pars = c("sigma_mu", "mu")),
  np = nuts_params(fitted_model_NB_hier_ncp)
)
bayesplot_grid(parcoord_with_divs, parcoord_no_divs, 
               ylim = c(-3, 3))
```

```{r samps-full-hier}
samps_NB_hier_ncp <- rstan::extract(fitted_model_NB_hier_ncp, pars = c('y_rep','inv_phi'))
```

The marginal plot, again.

```{r ppc-full-hier}
y_rep <- as.matrix(fitted_model_NB_hier_ncp, pars = "y_rep")
ppc_dens_overlay(stan_dat_hier$complaints, y_rep[1:200,])
```

This looks quite nice. If we've captured the building-level means well, then the
posterior distribution of means by building should match well with the observed
means of the quantity of building complaints by month.

```{r ppc-group_means-hier}
ppc_stat_grouped(
  y = stan_dat_hier$complaints, 
  yrep = y_rep,
  group = stan_dat_hier$building_idx, 
  stat = 'mean'
)
```

The building means appear to be well-captured by our model.

Predictions by number of traps:

```{r}
ppc_intervals(
  y = stan_dat_hier$complaints, 
  yrep = y_rep,
  x = stan_dat_hier$traps
) + 
  labs(x = "Number of traps", y = "Number of complaints")
```

Standardized residuals:
```{r}
mean_y_rep <- colMeans(y_rep)
mean_inv_phi <- mean(as.matrix(fitted_model_NB_hier_ncp, pars = "inv_phi"))
std_resid <- (stan_dat_hier$complaints - mean_y_rep) / sqrt(mean_y_rep + mean_y_rep^2*mean_inv_phi)
qplot(mean_y_rep, std_resid) + hline_at(2) + hline_at(-2)
```

Rootogram:
```{r}
ppc_rootogram(stan_dat_hier$complaints, yrep = y_rep)
```


### Varying intercepts _and_ varying slopes

We've gotten some new data that'll let us explore how to expand
the model a bit more with varying slopes along with varying intercepts
and time-varying intercepts.

```{r}
stan_dat_hier <- readRDS('data/pest_data_longer_stan_dat.RDS')
```

Perhaps if the levels of complaints
differ by building, the coefficient for the effect of traps on building does
too. We can add this to our model and observe the fit.

$$
\text{complaints}_{b,t} \sim \text{Neg-Binomial}(\lambda_{b,t}, \phi)  \\
\lambda_{b,t} = \exp{(\eta_{b,t})}\\
\eta_{b,t} = \mu_b + \kappa_b \, \texttt{traps}_{b,t} + \text{log_sq_foot}_b \\
\mu_b \sim \text{Normal}(\alpha + \texttt{building_data} \, \zeta, \sigma_{\mu}) \\
\kappa_b \sim \text{Normal}(\beta + \texttt{building_data} \, \gamma, \sigma_{\kappa})
$$

Let's compile the model.

```{r comp-NB-hier-slopes, cache=TRUE, results="hide", message=FALSE}
comp_model_NB_hier_slopes <- stan_model('stan_programs/hier_NB_regression_ncp_slopes_mod2.stan')
```

Fit the model to data and extract the posterior draws needed for 
our posterior predictive checks.

```{r run-NB-hier-slopes}
fitted_model_NB_hier_slopes <- 
  sampling(
    comp_model_NB_hier_slopes,
    data = stan_dat_hier,
    chains = 4, cores = 4,
    control = list(adapt_delta = 0.95)
  )
```

To see if the model infers building-to-building differences in, we can plot a
histogram of our marginal posterior distribution for `sigma_kappa`.

```{r}
mcmc_hist(
  as.matrix(fitted_model_NB_hier_slopes, pars = "sigma_kappa"), 
  binwidth = 0.005
)
```

```{r}
print(fitted_model_NB_hier_slopes, pars = c('kappa','beta','alpha','phi','sigma_mu','sigma_kappa','mu'))
```

```{r}
mcmc_hist(
  as.matrix(fitted_model_NB_hier_slopes, pars = "beta"), 
  binwidth = 0.005
)
```

While the model can't specifically rule out zero from the posterior, it does have mass at small
non-zero numbers, so we should leave in the hierarchy over $\texttt{kappa}$. Plotting the marginal
data density again, we can see the model still looks well calibrated.

```{r ppc-full-hier-slopes}
y_rep <- as.matrix(fitted_model_NB_hier_slopes, pars = "y_rep")
ppc_dens_overlay(
  y = stan_dat_hier$complaints, 
  yrep = y_rep[1:200,]
)
```


## Time varying effects and structured priors

We haven't looked at how cockroach complaints change over time. Let's look
at whether there's any pattern over time.

```{r ppc-group_max-hier-slopes-mean-by-mo}
select_vec <- which(stan_dat_hier$mo_idx %in% 1:12)
ppc_stat_grouped(
  y = stan_dat_hier$complaints[select_vec], 
  yrep = y_rep[,select_vec], 
  group = stan_dat_hier$mo_idx[select_vec], 
  stat = 'mean'
) + xlim(0, 11)
```

We might augment our model with a log-additive monthly effect, $\texttt{mo}_t$. 

$$
\eta_{b,t} = \mu_b + \kappa_b \, \texttt{traps}_{b,t} + \texttt{mo}_t + \text{log_sq_foot}_b
$$

We have complete freedom over how to specify the prior for $\texttt{mo}_t$. There are 
several competing factors for how the number of complaints might change over time. It makes
sense that there might be more roaches in the environment during the summer, but we might 
also expect that there is more roach control in the summer as well. Given that we're 
modeling complaints, maybe after the first sighting of roaches in a building, residents are
more vigilant, and thus complaints of roaches would increase.

This can be a motivation for using an autoregressive prior for our monthly effects. The model
is as follows:

$$
\texttt{mo}_t \sim \text{Normal}(\rho \, \texttt{mo}_{t-1}, \sigma_\texttt{mo}) \\
\equiv \\
\texttt{mo}_t = \rho \, \texttt{mo}_{t-1} +\epsilon_t , \quad \epsilon_t \sim \text{Normal}(0, \sigma_\texttt{mo}) \\
\quad \rho \in [-1,1]
$$

This equation says that the monthly effect in month $t$ is directly related to
the last month's monthly effect. Given the description of the process above,
it seems like there could be either positive or negative associations between
the months, but there should be a bit more weight placed on positive 
$\rho$s, so we'll put an informative prior that pushes the parameter $\rho$ towards
0.5.

Before we write our prior, however, we have a problem: Stan doesn't implement any densities
that have support on $[-1,1]$. We can use variable transformation of a raw variable
defined on $[0,1]$ to to give us a density on $[-1,1]$. Specifically,

$$
\rho_{\text{raw}} \in [0, 1] \\
\rho = 2 \times \rho_{\text{raw}} - 1
$$

Then we can put a beta prior on $\rho_\text{raw}$ to push our estimate towards 0.5.

One further wrinkle is that we have a prior for $\texttt{mo}_t$ that depends on
$\texttt{mo}_{t-1}$. What should we do about the prior for $\texttt{mo}_1$,
which doesn't have a previous time period in the data? We need to work out the
marginal distribution of the first observation. We can use the fact that AR
models are stationary, so
$\text{Var}(\texttt{mo}_t) = \text{Var}(\texttt{mo}_{t-1})$
and $\mathbb{E}(\texttt{mo}_t) = \mathbb{E}(\texttt{mo}_{t-1})$ for all $t$.

First we derive the marginal variance of $\texttt{mo}_{1}$:

$$
\text{Var}(\texttt{mo}_1) = \text{Var}(\rho \texttt{mo}_{0} + \epsilon_1)  \\
\text{Var}(\texttt{mo}_1) = \text{Var}(\rho \texttt{mo}_{0}) + \text{Var}(\epsilon_1)
$$
where the second line holds by independence of $\epsilon_1$ and $\epsilon_{0})$.
Then, using the fact that $Var(cX) = c^2Var(X)$ for a constant $c$ and the fact
that, by stationarity, $\textrm{Var}(\texttt{mo}_{0}) = \textrm{Var}(\texttt{mo}_{1})$, 
we then obtain:

$$
\text{Var}(\texttt{mo}_1)= \rho^2 \text{Var}( \texttt{mo}_{1})  + \sigma_\texttt{mo}^2 \\
\text{Var}(\texttt{mo}_1) = \frac{\sigma_\texttt{mo}^2}{1 - \rho^2}
$$

For the mean of $\texttt{mo}_1$ things are a bit simpler:

$$
\mathbb{E}(\texttt{mo}_1) = \mathbb{E}(\rho \, \texttt{mo}_{0} + \epsilon_t) \\
\mathbb{E}(\texttt{mo}_1) = \mathbb{E}(\rho \, \texttt{mo}_{0}) + \mathbb{E}(\epsilon_t) \\
$$

Since $\mathbb{E}(\epsilon_t) = 0$ by assumption we have 

$$
\mathbb{E}(\texttt{mo}_1) = \mathbb{E}(\rho \, \texttt{mo}_{0})  + 0\\
\mathbb{E}(\texttt{mo}_1) = \rho \, \mathbb{E}(\texttt{mo}_{1}) \\
\mathbb{E}(\texttt{mo}_1) - \rho \mathbb{E}(\texttt{mo}_1) = 0  \\
\mathbb{E}(\texttt{mo}_1) = 0/(1 - \rho)  
$$

which for $\rho \neq 1$ yields $\mathbb{E}(\texttt{mo}_{1}) = 0$. 

We now have the mean and variance for $\texttt{mo}_{1}$ so we can write down 
the full AR(1) process:

$$
\texttt{mo}_1 \sim \text{Normal}\left(0, \frac{\sigma_\texttt{mo}}{\sqrt{1 - \rho^2}}\right) \\
\texttt{mo}_t \sim \text{Normal}\left(\rho \, \texttt{mo}_{t-1}, \sigma_\texttt{mo}\right) \forall t > 1
$$

```{r comp-NB-hier-mos, cache=TRUE, results="hide", message=FALSE}
comp_model_NB_hier_mos <- stan_model('stan_programs/hier_NB_regression_ncp_slopes_mod_mos2.stan')
```

```{r run-NB-hier-slopes-mos}
fitted_model_NB_hier_mos <- sampling(comp_model_NB_hier_mos, data = stan_dat_hier, chains = 4, cores = 4, control = list(adapt_delta = 0.9))
```

In the interest of brevity, we won't go on expanding the model, though we certainly could.
What other information would help us understand the data generating process better? What
other aspects of the data generating process might we want to capture that we're not 
capturing now?

As usual, we run through our posterior predictive checks.

```{r ppc-full-hier-mos}
y_rep <- as.matrix(fitted_model_NB_hier_mos, pars = "y_rep")
ppc_dens_overlay(
  y = stan_dat_hier$complaints, 
  yrep = y_rep[1:200,]
)
```

```{r}
select_vec <- which(stan_dat_hier$mo_idx %in% 1:12)
ppc_stat_grouped(
  y = stan_dat_hier$complaints[select_vec], 
  yrep = y_rep[,select_vec], 
  group = stan_dat_hier$mo_idx[select_vec], 
  stat = 'mean'
) 
```

As we can see, our monthly random intercept has captured a monthly pattern
across all the buildings. We can also compare the prior and posterior for the
autoregressive parameter to see how much we've learned. Here are two different
ways of comparing the prior and posterior visually:

```{r}
# 1) compare draws from prior and draws from posterior
rho_draws <- cbind(
  2 * rbeta(4000, 10, 5) - 1, # draw from prior
  as.matrix(fitted_model_NB_hier_mos, pars = "rho")
)
colnames(rho_draws) <- c("prior", "posterior")
mcmc_hist(rho_draws, freq = FALSE, binwidth = 0.025,
          facet_args = list(nrow = 2)) + xlim(-1, 1)


# 2) overlay prior density curve on posterior draws
gen_rho_prior <- function(x) {
  alpha <- 10; beta <- 5
  a <- -1; c <- 1
  lp <- (alpha - 1) * log(x - a) + 
        (beta - 1) * log(c - x) - 
        (alpha + beta - 1) * log(c - a) - 
         lbeta(alpha, beta)
  return(exp(lp))
}
mcmc_hist(as.matrix(fitted_model_NB_hier_mos, pars = "rho"), 
          freq = FALSE, binwidth = 0.01) +
  overlay_function(fun = gen_rho_prior) + 
  xlim(-1,1)
```
```{r}
print(fitted_model_NB_hier_mos, pars = c('rho','sigma_mu','sigma_kappa','gamma'))
```

```{r}
ppc_intervals(
  y = stan_dat_hier$complaints, 
  yrep = y_rep,
  x = stan_dat_hier$traps
) + 
  labs(x = "Number of traps", y = "Number of complaints")
```

It looks as if our model finally generates a reasonable posterior predictive
distribution for all numbers of traps, and appropriately captures the tails of the 
data generating process.


## Using our model: Cost forecasts

Our model seems to be fitting well, so now we will go ahead and use the model to
help us make a decision about how many traps to put in our buildings. We'll make
a forecast for 6 months forward.

```{r comp-rev, cache=TRUE, results="hide", message=FALSE}
comp_rev <- stan_model('stan_programs/hier_NB_regression_ncp_slopes_mod_mos_predict.stan')
print(comp_rev)
```

An important input to the revenue model is how much revenue is lost due to each
complaint. The client has a policy that for every 10 complaints, they'll call an
exterminator costing the client \$100, so that'll amount to \$10 per complaint.

```{r run-NB-hier-rev}
rev_model <- sampling(comp_rev, data = stan_dat_hier, cores = 4,
                      control = list(adapt_delta = 0.9))
```

Below we've generated our revenue curves for the buildings. These charts will
give us precise quantification of our uncertainty around our revenue projections
at any number of traps for each building.

A key input to our analysis will be the cost of installing bait stations. We're 
simulating the number of complaints we receive over the course of a year, so we 
need to understand the cost associated with maintaining each bait station over 
the course of a year. There's the cost attributed to the raw bait station, which
is the plastic housing and the bait material, a peanut-buttery substance that's 
injected with insecticide. The cost of maintaining one bait station for a year 
plus monthly replenishment of the bait material is about \$20. 

```{r materials-cost-formulation}
N_traps <- 20
costs <- 10 * (0:N_traps) 
```

We'll also need labor for maintaining the bait stations, which need to be
serviced every two months. If there are fewer than five traps, our in-house
maintenance staff can manage the stations (about one hour of work every two
months at \$20/hour), but above five traps we need to hire outside pest control
to help out. They're a bit more expensive, so we've put their cost at \$30 /
hour. Each five traps should require an extra person-hour of work, so that's
factored in as well. The marginal person-person hours above five traps are at
the higher pest-control-labor rate.

```{r labor-cost-formulation}
N_months_forward <- 12
N_months_labor <- N_months_forward / 2
hourly_rate_low <- 20
hourly_rate_high <- 30
costs <- costs + 
  (0:N_traps < 5 & 0:N_traps > 0) * (N_months_labor * hourly_rate_low) + 
  (0:N_traps >= 5 & 0:N_traps < 10) * (N_months_labor * (hourly_rate_low + 1 * hourly_rate_high)) + 
  (0:N_traps >= 10 & 0:N_traps < 15) * (N_months_labor * (hourly_rate_low + 2 * hourly_rate_high)) + 
  (0:N_traps >= 15) * (N_months_labor * (hourly_rate_low + 3 * hourly_rate_high))
```

We can now plot curves with number of traps on the x-axis and profit/loss
forecasts and uncertainty intervals on the y-axis.

```{r rev-curves, fig.width = 6, fig.height = 8}
# extract as a list for convenience below
samps_rev <- rstan::extract(rev_model)

# total and mean profit
tot_profit <- sweep(samps_rev$rev_, 3, STATS = costs, FUN = '-')
mean_profit <- t(apply(tot_profit, c(2, 3), median))

# lower and upper ends of 50% central interval
lower_profit <- t(apply(tot_profit, c(2, 3), quantile, 0.25))
upper_profit <- t(apply(tot_profit, c(2, 3), quantile, 0.75))

profit_df <-
  data.frame(
    profit = as.vector(mean_profit),
    lower = as.vector(lower_profit),
    upper = as.vector(upper_profit),
    traps = rep(0:N_traps, times = N_buildings),
    building = rep(1:N_buildings, each = N_traps + 1)
  )
ggplot(data = profit_df, aes(x = traps, y = profit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line() +
  facet_wrap(~ building, scales = 'free_y', ncol = 2)
```

We can can see that the optimal number of bait stations differs by building.


<br>

Left as an exercise for the reader: 

* How would we build a revenue curve for a new building?

* Let's say our utility function is revenue. If we wanted to maximize expected
revenue, we can take expectations at each station count for each building, and
choose the trap numbers that maximizes expected revenue. This will be called a
maximum revenue strategy. How can we generate the distribution of portfolio
revenue (i.e. the sum of revenue across all the buildings) under the maximum
revenue strategy from the the draws of `rev_pred` we already have?


### Gaussian process instead of AR(1)
```{r comp-NB-hier-mos, cache=TRUE, results="hide", message=FALSE}
comp_model_NB_hier_gp <- stan_model('stan_programs/hier_NB_regression_ncp_slopes_mod_mos_gp.stan')
```

```{r run-NB-hier-slopes-mos}
fitted_model_NB_hier_gp <- sampling(comp_model_NB_hier_gp, data = stan_dat_hier, chains = 4, cores = 4, control = list(adapt_delta = 0.9))
```

# Useful resources: 

[ADD MORE TO THIS]

* [The Stan Forums](http://discourse.mc-stan.org/)
* [Stan Language Documentation](http://mc-stan.org/users/documentation/)
* [Stan Case Studies](http://mc-stan.org/users/documentation/case-studies)
* **rstan** [vignettes](http://mc-stan.org/rstan/articles)
* **bayesplot** tutorial [vignettes](http://mc-stan.org/bayesplot/articles)
* **loo** tutorial [vignettes](http://mc-stan.org/loo/articles)
* [Interactive MCMC demos](https://chi-feng.github.io/mcmc-demo/)
* Paper: [_A Conceptual Introduction to Hamiltonian Monte Carlo_](https://arxiv.org/abs/1701.02434)
* Paper: [_Visualization in Bayesian Workflow_](https://github.com/jgabry/bayes-vis-paper#readme)
* Link to prophet package
